---
title: "Project2_5300"
author: "LAVANYA B"
output: html_document
---


#1. Does the average gain differ for flights that departed late versus those that did not? What about for flights that departed more than 30 minutes late?

```{r}

# Load necessary libraries
library(nycflights13)
library(tidyverse)
library(ggplot2)

# Load flight data
flights_data <- nycflights13::flights

# Filter for United Airlines (UA) flights
ua_flights_data <- flights_data %>%
  filter(carrier == "UA")

# Create a new variable for net gain
ua_flights_data <- ua_flights_data %>%
  mutate(net_gain = dep_delay - arr_delay)

# Perform t-test for flights that departed late versus those that did not
t_test_result_late_vs_not_late <- t.test(net_gain ~ (dep_delay > 0), data = ua_flights_data)

# Display the t-test result
print("T-Test for Average Gain for Flights that Departed Late versus Not Late:")
print(t_test_result_late_vs_not_late)

# Perform t-test for flights that departed more than 30 minutes late versus those that did not
t_test_result_late_more_than_30_vs_not_late <- t.test(net_gain ~ (dep_delay > 30), data = ua_flights_data)

# Display the t-test result
print("T-Test for Average Gain for Flights that Departed More than 30 Minutes Late versus Not Late:")
print(t_test_result_late_more_than_30_vs_not_late)

# Create boxplot for net gain comparing flights that departed late versus not late
ggplot(ua_flights_data, aes(x = ifelse(dep_delay > 0, "Late", "On Time"), y = net_gain, fill = ifelse(dep_delay > 0, "Late", "On Time"))) +
  geom_boxplot() +
  labs(title = "Boxplot of Net Gain for Flights that Departed Late Vs On Time",
       x = "Departure Delay Status",
       y = "Net Gain") +
  scale_fill_discrete(name = "Departure Delay") +
  theme_minimal()

# Plot boxplot for net gain comparing flights that departed more than 30 minutes late versus not late
ggplot(ua_flights_data, aes(x = ifelse(dep_delay > 30, "Late", "On Time"), y = net_gain, fill = ifelse(dep_delay > 30, "Late", "On Time"))) +
  geom_boxplot() +
  labs(title = "Boxplot of Net Gain for Flights that Departed More than 30 Minutes Late Vs On Time",
       x = "Departure Delay Status",
       y = "Net Gain") +
  scale_fill_discrete(name = "Departure Delay") +
  theme_minimal()



```


#2. What are the five most common destination airports for United Airlines flights from New York City? Describe the distribution and the average gain for each of these five airports.

```{r}

library(nycflights13)

# Filter United Airlines (UA) flights from New York City
ua_flights <- filter(flights, carrier == "UA", origin %in% c("JFK", "LGA", "EWR"))

# Calculate net gain for each flight
ua_flights <- mutate(ua_flights, net_gain = dep_delay - arr_delay)

# Identify the five most common destination airports
top_destinations <- ua_flights %>%
  group_by(dest) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5)
top_destinations

```
```{r}
# Describe the distribution and average gain for each of these five airports
summary_stats <- ua_flights %>%
  filter(dest %in% top_destinations$dest) %>%
  group_by(dest) %>%
  summarise(
    avg_gain = mean(net_gain, na.rm = TRUE),
    median_gain = median(net_gain, na.rm = TRUE),
    sd_gain = sd(net_gain, na.rm = TRUE)
    
  )

print(summary_stats)
```

```{r}
# T-test for each destination airport
t_test_results <- ua_flights %>%
  filter(dest %in% top_destinations$dest) %>%
  group_by(dest) %>%
  summarise(
    t_statistic = t.test(net_gain)$statistic,
    p_value = t.test(net_gain)$p.value
  )

# Confidence intervals for each destination airport
ci_results <- ua_flights %>%
  filter(dest %in% top_destinations$dest) %>%
  group_by(dest) %>%
  summarise(
    avg_gain = mean(net_gain, na.rm = TRUE),
    ci_lower = t.test(net_gain)$conf.int[1],
    ci_upper = t.test(net_gain)$conf.int[2]
  )

# Combine t-test and confidence interval results
combined_results <- left_join(t_test_results, ci_results, by = "dest")

print(combined_results)
```

```{r}
# Create a bar graph
ggplot(ci_results, aes(x = dest, y = avg_gain, fill = dest)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(0.9), width = 0.25) +
  labs(title = "Average Net Gain and Confidence Interval by Destination Airport",
       x = "Destination Airport",
       y = "Average Net Gain") +
  theme_minimal()
```


#3. Another common measure of interest, in addition to total gain, is the gain relative to the duration of the flight. Calculate the gain per hour by dividing the total gain by the duration in hours of each flight. Does the average gain per hour differ for flights that departed late versus those that did not? What about for flights that departed more than 30 minutes late?


```{r}
library(nycflights13)
library(dplyr)
library(ggplot2)

# Filter United Airlines (UA) flights from New York City
ua_flights <- filter(flights, carrier == "UA", origin %in% c("JFK", "LGA", "EWR"))

# Calculate net gain for each flight
ua_flights <- mutate(ua_flights, net_gain = dep_delay - arr_delay)

# Calculate gain per hour
ua_flights <- mutate(ua_flights, gain_per_hour = net_gain / (air_time / 60))

# Create a new variable for late departure (more than 30 minutes late)
ua_flights <- mutate(ua_flights, late_departure = ifelse(dep_delay > 30, "Late Departure", "On Time"))

# Perform t-test to compare average gain per hour for late departures vs. on-time departures
t_test_result <- t.test(gain_per_hour ~ late_departure, data = ua_flights)

# Print the t-test result
print(t_test_result)

# Create a box plot to compare gain per hour for late departures vs. on-time departures
ggplot(ua_flights, aes(x = late_departure, y = gain_per_hour, fill = late_departure)) +
  geom_boxplot() +
  geom_text(aes(label = paste("p =", format(t_test_result$p.value, digits = 2))), 
            x = 1.5, y = max(ua_flights$gain_per_hour), hjust = 0.5, vjust = -0.5, color = "black") +
  labs(title = "Box Plot of Gain Per Hour by Departure Status",
       x = "Departure Status",
       y = "Gain Per Hour (minutes)") +
  scale_fill_manual(values = c("lightblue", "lightcoral")) +
  theme_minimal()

```

#4. Does the average gain per hour differ for longer flights versus shorter flights?

```{r}
# Load necessary libraries
library(nycflights13)
library(dplyr)

# Filter United Airlines (UA) flights from New York City
ua_flights <- filter(flights, carrier == "UA", origin %in% c("JFK", "LGA", "EWR"))

# Calculate net gain for each flight
ua_flights <- mutate(ua_flights, net_gain = dep_delay - arr_delay)

# Calculate gain per hour
ua_flights <- mutate(ua_flights, gain_per_hour = net_gain / (air_time / 60))

# Define a threshold for flight duration (e.g., 180 minutes, adjust as needed)
duration_threshold <- 180

# Create new variables for longer and shorter flights
ua_flights <- mutate(ua_flights, 
                     longer_flight = air_time > duration_threshold,
                     shorter_flight = air_time <= duration_threshold)

# Perform t-test to compare average gain per hour for longer flights vs. shorter flights
t_test_longer <- t.test(gain_per_hour ~ longer_flight, data = ua_flights)

# Perform t-test to compare average gain per hour for shorter flights vs. longer flights
t_test_shorter <- t.test(gain_per_hour ~ shorter_flight, data = ua_flights)


print(t_test_longer)

print(t_test_shorter)

```

```{r}
# Create a box plot to compare gain per hour for longer flights vs. shorter flights
ggplot(ua_flights, aes(x = factor(longer_flight), y = gain_per_hour, fill = factor(longer_flight))) +
  geom_boxplot() +
  labs(title = "Box Plot of Gain Per Hour by Flight Duration",
       x = "Flight Duration",
       y = "Gain Per Hour (minutes)") +
  scale_fill_manual(values = c("lightblue", "lightcoral"), labels = c("Shorter Flights", "Longer Flights")) +
  theme_minimal()
```






















